\section{Proposed Contributions}
\label{sec:proposed_approach}
We propose three further developments that could contribute to enhance the
effectiveness, the precision and the generality of \textsc{Phoenix}:
\begin{itemize}
  \item Incremental Clustering
  \item Multilingual/Multicharset Support
  \item Outliers detection
\end{itemize}
In the next paragraphs we shall discuss them thoroughly.

\subsection{Incremental Clustering} % (fold)
\label{sub:incremental_clustering}
So far \textsc{Phoenix} allows a \emph{batch} approach. The data must be crunched as
a whole, if new domain names are added to the dataset it is not possible to incrementally
update the obtained results. We want to integrate in our tool the techniques proposed in
\cite{Ning2010} and \cite{Chi2009} for incremental clustering, and we want to
substitute the Normalized Cuts technique for spectral clustering \cite{Shi2000} with
the more recent SPAN method developed by Shu \emph{et al.} \cite{Shu2011}.

% subsection incremental_clustering (end)

\subsection{Multilingual/Multicharset Support} % (fold)
\label{sub:multilingual_multicharset_support}
The linguistic features of \textsc{Phoenix} were extracted using the English
dictionary. We want to investigate the presence of other languages in malicious
domain names, e.g. Swedish or French. If we will find evidence of a substantial
presence of such languages, it is our intention to develop a multilingual module,
able to select correctly the dictionary to be used.

Moreover it is recently possible to use non-ASCII domain names, e.g.
www.$\pi$.com. This is an uncharted area that we want to explore as it might lead
to interesting results.

% subsection multilingual_multicharset_support (end)

\subsection{Outliers Detection} % (fold)
\label{sub:outliers_detection}
We want to investigate the possibility of using different clustering algorithms
(other than DBSCAN) to see if it leads to better performances. Especially it is of
our interest to search for an efficient technique for outliers detection purposes.

% subsection outliers_detection (end)

% section proposed_approach (end)
